---
title: "A4: Prison Data"
author: "Manav Agarwal"
date: '2022-11-21'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
[TODO: Include a short abstract (about 50 words)]

# Test (TODO: Remove before submission!!)
```{r echo=FALSE}
q1_string <- test_query1()
num <- 5
```
As you know, this is a markdown report. The [R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is **useful!**

* Here's a test of a dynamic variable. Say **"`r q1_string`!!!** 
* Here's another test. A vector
of numbers: `r test_query2(num)`.

Recall that _{r echo=TRUE}_ can be used to display the code, 
which is very useful when developing and debugging a report. Here's an example: 

```{r echo=TRUE}
num <- 5
sum <- sum(test_query2(num))
print(sum)
```

## Introduction 
The data from Vera Institute deals with incarceration data from 1970 - 2018 classified by various racial demographics. This data can help us showcase which of these demographics are minoritized, and disproportionately affected by, compared to those that have relative power in this society. For this specific assignment, I intend to analyzing the dataset relating to the Asian American and Pacific Islanders population. As observed in the news, there was an increase on attacks and disproportionate violence and racism against this demographic, during the peak of the COVID-19 pandemic, and through this assignment I hope to track this over a period of time. The direct stakeholders of this dataset are the people whose information is contained in this. The indirect stakeholders are the entire current and future populations of these various groups classified in this dataset.
I intend on analyzing the trends in the crimes and prison rate of African American demographics. 
## Data Summary 
[TODO]

[TODO ... other sections ... See Canvas for details]

_All the very best with your work!_

